Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-12 aistats.tex
Start formatting figures

@@

\begin{figure*}[ht]
	\centering
	\includegraphics[width=\textwidth]{{figs/biasvariance/new_linear_ratio=0.8_n0=10}.pdf}
	\caption{
	\textbf{Gaussian variance $\cN(0,\sigma^2)$ example. Left:} training curves and analytic upper bound.
	\textbf{Center:} bias-mixed-variance decomposition, using the arithmetic mean.
	\textbf{Right:} bias-variance decomposition, using the harmonic mean.
	}
	\label{fig:variance_decomposition}
\end{figure*}

\begin{figure*}[ht]
	\centering
	\includegraphics[width=\textwidth]{{figs/gaussians/new_linear_n0=1}.pdf}
	\caption{
	\textbf{Full gaussian} $\cN(m, \sigma^2)$ with $\meanp^*=(0, 1), \meanp_0 = (1,2)$ and $n_0=1$. \textbf{Left:} training curves (5,95) percentiles, average and asymptote.
	\textbf{Right:} bias-variance decomposition.
	}
	\label{fig:gaussian_decomposition}
\end{figure*}

@@

\begin{figure}[ht]
	\centering
\includegraphics[width=.4\textwidth]{fewsamples.pdf}
	\caption{Suboptimality of a Gaussian variance MLE against number of samples $n$. Bold curve is average over 100 trials,  dark shaded area is 90\% (dark) confidence interval, light shade is min-max interval.
		While the expected value is infinite for $n=1$ or $n=2$, it quickly matches the upper bound, and the $1/2n$ asymptote.
	}
	\label{fig:curves}
\end{figure}

@@

This inequality highlights a clear variance-bias decomposition.
In particular, there is no bias term when $\frac{\m_0}{\m^*} =1 + \inv{n_0} $, which happens when the prior is slightly larger than the ground truth.  For instance, when $n_0=1$, it encourages us to set $\m_0 = 2 \m^*$.
Remark that the variance term is not asymptotically tight as the log-inequality we used \eqref{eq:log_bound} is not quadratically tight around $1$. We are basically losing a factor 2 compared to $1/2n$.

@@

\newpage~\null\\
\FloatBarrier

@@

\subsection{Relative Smoothness}
Mirror descent (MD)\footnote{
Also known as
Bregman (proximal) gradient, relative gradient descent or NoLips.
}
\citep{nemirovski1983problem,beck2003mirror}
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citet{nemirovski2009robust,ghadimi2012optimal}  \TODO{verify what these citations are talking about} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2 \& Th.6.3]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

@@

\begin{table}[t]
\begingroup
\renewcommand*{\arraystretch}{1.25}%
\newcommand*{\greencmark}{\textcolor{Green}{\cmark}}
\newcommand*{\redxmark}{\textcolor{Red}{\xmark}}
\centering
\begin{tabular}{lccc}
%{p{.15\textwidth}m{.05\textwidth}m{.08\textwidth}m{.1\textwidth}}
\toprule
Boundedness & Barrier &  $\lr_n \sim \inv{n}$ & Last iterate \\
\midrule
%Strongly convex and bounded variance\newline
%\citep[e.g.,][]{??}
%&
%$
%\cB_A(\theta,\theta') \geq \frac{1}{2}\norm{\theta-\theta'}^2,
%\quad
%\expect[x]{\norm{\m - x}}^2\leq\sigma^2
%$
%\\
Variance on all $\Theta$ % \newline \citep{hanzely2018fastest}
& \redxmark & \greencmark  & \redxmark
\\
Variance at $\theta_*$ %\newline \citep{dragomir2021fast}
& \redxmark & \greencmark  & \greencmark
\\
Optimality gap %\newline \citep{dorazio2021stochastic}
& \greencmark & \redxmark & \greencmark
\\
\bottomrule
\end{tabular}
\caption{
Summary of results about SMD with relative smoothness and relative strong convexity assumptions. 
Barrier tells whether the assumption can hold with barrier objectives.
$\lr_n \sim \inv{n}$ says whether the convergence rate hold with this step-size scheduling, and the last column tells whether the rate applies to the last iterate or to some average iterate.
}
\label{tbl:assumptions}
\endgroup
\end{table}

>>>>

\begin{figure}[t]
	\centering
\includegraphics[width=.4\textwidth]{fewsamples.pdf}
	\caption{Suboptimality of a Gaussian variance MLE against number of samples $n$. Bold curve is average over 100 trials,  dark shaded area is 90\% (dark) confidence interval, light shade is min-max interval.
		While the expected value is infinite for $n=1$ or $n=2$, it quickly matches the upper bound, and the $1/2n$ asymptote.
	}
	\label{fig:curves}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{{figs/gaussian_variance.pdf}}
	\caption{
	\textbf{Gaussian variance $\cN(0,\sigma^2)$ example. Left:} training curves and analytic upper bound.
	\textbf{Center:} bias-mixed-variance decomposition, using the arithmetic mean.
	\textbf{Right:} bias-variance decomposition, using the harmonic mean.
	}
	\label{fig:variance_decomposition}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=.4\textwidth]{{figs/gaussians/new_linear_n0=1}.pdf}
	\caption{
	\textbf{Full gaussian} $\cN(m, \sigma^2)$ with $\meanp^*=(0, 1), \meanp_0 = (1,2)$ and $n_0=1$. \textbf{Left:} training curves (5,95) percentiles, average and asymptote.
	\textbf{Right:} bias-variance decomposition.
	}
	\label{fig:gaussian_decomposition}
\end{figure}

@@

@@

This inequality highlights a clear variance-bias decomposition.
In particular, there is no bias term when $\frac{\m_0}{\m^*} =1 + \inv{n_0} $, which happens when the prior is slightly larger than the ground truth.  For instance, when $n_0=1$, it encourages us to set $\m_0 = 2 \m^*$. This correlates well with numerical observations.
Remark that the variance term is not asymptotically tight as the log-inequality we used \eqref{eq:log_bound} is not quadratically tight around $1$. We are basically losing a factor 2 compared to $1/2n$.

@@

@@

\subsection{Relative Smoothness}
Mirror descent (MD)\footnote{
Also known as
Bregman (proximal) gradient, relative gradient descent or NoLips.
}
\citep{nemirovski1983problem,beck2003mirror}
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citep{nemirovski2009robust,ghadimi2012optimal} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2 \& Th.6.3]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

@@

\begin{table}[t]
\begingroup
\renewcommand*{\arraystretch}{1.25}%
\newcommand*{\greencmark}{\textcolor{Green}{\cmark}}
\newcommand*{\redxmark}{\textcolor{Red}{\xmark}}
\centering
\begin{tabular}{lccc}
%{p{.15\textwidth}m{.05\textwidth}m{.08\textwidth}m{.1\textwidth}}
\toprule
Boundedness & Barriers &  $\lr_n \sim \inv{n}$ & Last iterate \\
\midrule
%Strongly convex and bounded variance\newline
%\citep[e.g.,][]{??}
%&
%$
%\cB_A(\theta,\theta') \geq \frac{1}{2}\norm{\theta-\theta'}^2,
%\quad
%\expect[x]{\norm{\m - x}}^2\leq\sigma^2
%$
%\\
Variance on all $\Theta$ % \newline \citep{hanzely2018fastest}
& \redxmark & \greencmark  & \redxmark
\\
Variance at $\theta_*$ %\newline \citep{dragomir2021fast}
& \redxmark & \greencmark  & \greencmark
\\
Optimality gap %\newline \citep{dorazio2021stochastic}
& \greencmark & \redxmark & \greencmark
\\
\bottomrule
\end{tabular}
\caption{
Summary of results about SMD with relative smoothness and relative strong convexity assumptions. 
Barrier tells whether the assumption can hold with barrier objectives.
$\lr_n \sim \inv{n}$ says whether the convergence rate hold with this step-size scheduling, and the last column tells whether the rate applies to the last iterate or to some average iterate.
}
\label{tbl:assumptions}
\endgroup
\end{table}
